---
title: "Housing_Price"
author: "Yeshwanthi"
date: "6/6/2020"
output:
  html_document: default
  pdf_document: default
---

##Data Loading##

Loading crime data


```{r}
crimedata <- read.table("C:/Users/Deepak Yesh/Documents/Final-Project-DA/crime.csv", header=TRUE,sep=",")
```
Loading distance data
```{r}
distancedata <- read.table("C:/Users/Deepak Yesh/Documents/Final-Project-DA/amazon.csv",header=TRUE,sep=",")
```
Loading housing data
```{r}

housingdata <- read.table("C:/Users/Deepak Yesh/Documents/Final-Project-DA/initial-housingdata.csv",header=TRUE,sep=",")
```

##Data Merging##

Merging crime and distance data and storing in distancecrime dataframe
```{r}
distancecrime <- merge(distancedata,crimedata, by ="zipcode")
```

Merging distance crimem data with housing data and storing in fulldata data frame
```{r}
fulldata <- merge(distancecrime,housingdata, by="zipcode")
```

##Data Cleaning##

Stripping out comma and $ form Income

```{r}
fulldata$Income= as.numeric(gsub("\\$|,","",as.character(fulldata$Income)))
```
Stripping out comma from population

```{r}
fulldata$population= as.numeric(gsub("\\$|,","",as.character(fulldata$population)))
```

##Derived variables##

Deriving year from date

```{r}
fulldata$year<-substr(fulldata$date,1,4)
```

Deriving new living area square feet based on year of renovation

```{r}
fulldata$newsqft_living <- ifelse(fulldata$year==2015,fulldata$sqft_living15,fulldata$sqft_living)
```

Deriving new lot area square feet based on year of renovation

```{r}
fulldata$newsqft_lot <- ifelse(fulldata$year==2015,fulldata$sqft_lot15,fulldata$sqft_lot)
```

Deriving crime rate based on population

```{r}
fulldata$crimerate <-fulldata$crime.count/fulldata$population
fulldata$crimerate <- round(fulldata$crimerate,2)
```

##Analysing data issue problem##

```{r}
Rooms_Issue = fulldata$bedrooms
Rooms_Issue.freq = table(Rooms_Issue)
Rooms_Issue.freq
```

Removed data entry issue from the data set 
Bed room is 33 but square feet is very less

```{r}
fulldata[which(fulldata$bedrooms == 33),]
fulldata <- fulldata[-13247,]
```

##Exploratory Analysis for X Variables##

Water front

```{r}
library(plyr)
library(MASS) 
waterfrong = fulldata$waterfront
waterfrong.freq = table(waterfrong)
colors = c("antiquewhite", "yellow", "green", "violet", 
           "orange", "blue", "pink", "cyan") 
barplot((waterfrong.freq),main="Water Front House Freq Count",
        xlab="'O -- No WaterFront' '1 -- WaterFront' ",
        ylab="Freq Counts",col=colors)
```
Grade
```{r}
Gradeb = fulldata$grade
Gradeb.freq = table(Gradeb)
Gradeb.freq
colors = c("antiquewhite", "yellow", "cadetblue1", "aquamarine2", 
           "orange", "coral2", "pink", "cyan") 
barplot((Gradeb.freq),main="Grade Disturbution",
        xlab="Grade of the House based on King County Common Scale",
      ylab="Freq Counts",col=colors)
```

View
```{r}
View1 = fulldata$view
View1.freq = table(View1)
View1.freq
colors = c("antiquewhite", "yellow", "cadetblue1", "aquamarine2", 
           "orange") 
barplot((View1.freq),main="View Disturbution",
        xlab="Number of views for a house",
        ylab="Freq Counts",col=colors)
```

Condition
```{r}
condition1 = fulldata$condition
condition1.freq = table(condition1)
condition1.freq
colors = c("antiquewhite", "yellow", "cadetblue1", "aquamarine2", 
           "orange") 
barplot((condition1.freq),main="Condition Disturbution",
        xlab="Condition a house",
        ylab="Freq Counts",col=colors)

```


##Analysing the frequency of ordinal variables##

```{r}
library(car)
count(mtcars,'fulldata$view')
count(mtcars,'fulldata$waterfront')
count(mtcars,'fulldata$condition')
count(mtcars,'fulldata$grade')
count(mtcars,'fulldata$floors')
count(mtcars,'fulldata$bedroom')
```

##Exploratory Anlysis for y variable price##

```{r}

x<-fulldata$price
h<-hist(x, breaks=10, col="Blue", xlab="Price", 
        main="Histogram for Price") 
xfit<-seq(min(x),max(x),length=40) 
yfit<-dnorm(xfit,mean=mean(x),sd=sd(x)) 
yfit <- yfit*diff(h$mids[1:2])*length(x) 
lines(xfit, yfit, col="Red", lwd=2)
```

Applying log transformations on y variable

```{r}
x<-log(fulldata$price)
h<-hist(x, breaks=10, col="Blue", xlab="Price", 
        main="Histogram for Price") 
xfit<-seq(min(x),max(x),length=40) 
yfit<-dnorm(xfit,mean=mean(x),sd=sd(x)) 
yfit <- yfit*diff(h$mids[1:2])*length(x) 
lines(xfit, yfit, col="Red", lwd=2)
```
##Extracting the required variables to fulldata##

```{r}
data1<-fulldata[c(2,3,13,14,15,18,19,20,21,22,31,33,34,35)]
```
##Pearson correlation matrix##

```{r}
library(corrplot)
c<-cor(data1)
corrplot(c,method="number")
```


##Scatter plot for each X and Variable ##
```{r}
library(car)
plot(log(data1$price),data1$grade)
plot(log(data1$price),data1$Distance)
plot(log(data1$price),data1$Income)
plot(log(data1$price),data1$bedrooms)
plot(log(data1$price),data1$bathrooms)
plot(log(data1$price),data1$floors)
plot(log(data1$price),data1$waterfront)
plot(log(data1$price),data1$view)
plot(log(data1$price),data1$condition)
plot(log(data1$price),data1$newsqft_living)
plot(log(data1$price),data1$newsqft_lot)
```

##Splitting Training and Testing Data##

```{r}

selectfulldata = sample(1:nrow(fulldata),0.80*nrow(fulldata));
train1.data = data1[selectfulldata,]; 
test1.data = data1[-selectfulldata,];
str(train1.data)

```

##Linear Regression Model##

Removed non significiant variables -condition,squareftlot,view,waterfront

```{r}
model1 <- lm(formula=log(price)~bathrooms+grade+bedrooms+floors
         +newsqft_living+crimerate+Distance,data=train1.data)
summary(model1)
```

Removing bedroms as it is not significant

```{r}
model2 <- lm(formula=log(price)~bathrooms+grade+floors
           +newsqft_living+crimerate+Distance,data=train1.data)
summary(model2)
```

##Models using step wise regression##

Backward stepwise regression

```{r}
step(model1,direction="backward",trace=T)
#Backward suggested to remove bedrooms
#This model is same as model3

```

Forward stepwise regression

```{r}
step(model1,direction="forward",trace=T)
#Forward did not suggest to remove any variables
#This model is same as model2
```

Both stepwise regression
```{r}
step(model1,direction="both",trace=T)
#Both stepwiseregression suggested to remove bedrooms
#This model is same as model3
```

##Best subset Regression ##

Assigning x and y variable to leaps functions and using stpewise regression 
with cp as metric for feature selection
```{r}
library(leaps)
leaps(x=train1.data[c(1,2,4,5,6,10,12,14)],y=train1.data[,3],
      names=names(train1.data)[c(1,2,4,5,6,10,12,14
)],method="Cp")
#This suggested variables sames as model2
```

Assigning x and y variable to leaps functions and using stpewise regression 
with adjusted R squared squared as metric for feature selection

```{r}
leaps(x=train1.data[c(1,4,5,6,10,12,14)],y=train1.data[,3],
      names=names(train1.data)[c(1,4,5,6,10,12,14
      )],method="adjr2")
#This is suggesting distance,grade and square feet living as variables
```

Building a new model with variables suggested using adjusted R squared metric

```{r}
model3 <- lm(formula=log(price)~grade+
           newsqft_living+Distance,data=train1.data)
summary(model3)
```

Assigning x and y variable to leaps functions and using stpewise regression 
with R squared as metric for feature selection

```{r}
leaps(x=train1.data[c(1,4,5,6,10,12,14)],y=train1.data[,3],
      names=names(train1.data)[c(1,4,5,6,10,12,14
      )],method="r2")
#This suggest variables similar to m4
```


##Multicollinearity##

```{r}
library(car)
#Testing multicollinearity for model 1
vif(model1)
#Result - No Multi Colliearity issues in model1
```
```{r}
#Testing multicollinearity for model 2
vif(model2)
#Result - No Multi Colliearity issues in model2
```

```{r}
#Testing multicollinearity for model 3
vif(model3)
#Result - No Multi Colliearity issues in model3
```

##Predicted Values and rmse##

Assigning price as y

```{r}
y=test1.data[,3]
```

predicting the price for test data using model1

```{r}
m1y1<-exp(predict.glm(model1,test1.data))
```

Calculating rmse for predicted and observed values using model1

```{r}
py=test1.data[,3]
rmsem1 = sqrt((py-m1y1)%*%(py-m1y1))/nrow(test1.data)
rmsem1
```

predicting the price for test data using model1

```{r}
m2y2<-exp(predict.glm(model2,test1.data))
```

Calculating rmse for predicted and observed values using model1

```{r}
py=test1.data[,3]
rmsem2 = sqrt((py-m2y2)%*%(py-m2y2))/nrow(test1.data)
rmsem2
```

predicting the price for test data using model1

```{r}
m3y3<-exp(predict.glm(model3,test1.data))
```

Calculating rmse for predicted and observed values using model1
```{r}
py=test1.data[,3]
rmsem3 = sqrt((py-m3y3)%*%(py-m3y3))/nrow(test1.data)
rmsem3
```

Model 1

```{r}
rmsem1
```

Model 2
```{r}
rmsem2
```

Model 3

```{r}
rmsem3
```


##Residuals Assumptions##

model1
Residual plot normality
```{r}
qqnorm(rstandard(model1))
qqline(rstandard(model1,col="red"))
```

plot between residual and predicted
```{r}
plot(fitted(model2),rstandard(model2),main="predicted vs residual plot")
abline(a=0, b=0, col='red') 
```

model2
Residual plot normality
```{r}
qqnorm(rstandard(model2))
qqline(rstandard(model2,col="red"))
```

plot between residual and predicted
```{r}
plot(fitted(model2),rstandard(model2),main="predicted vs residual plot")
abline(a=0, b=0, col='red')
```

model3
Residual plot normality
```{r}
qqnorm(rstandard(model3))
qqline(rstandard(model3,col="red"))
```

plot between residual and predicted
```{r}
plot(fitted(model3),rstandard(model3),main="predicted vs residual plot")
abline(a=0, b=0, col='red')
```

##Residual vs independent variables for model3##

plot for residual and sqft living variable

```{r}
plot(train1.data$newsqft_living,rstandard(model2),main="Square Feet vs Residual plot")
abline(a=0, b=0, col='red')
```


plot for residual and grade variable
```{r}

plot(train1.data$grade,rstandard(model2),main="Grade vs Residual plot")
abline(a=0, b=0, col='red')
```


plot for residual and bedrooms variable
```{r}
plot(train1.data$bathrooms,rstandard(model2),main="Bathrooms vs Residual plot")
abline(a=0, b=0, col='red')
```

plot for residual and crime variable
```{r}
plot(train1.data$crimerate,rstandard(model2),main="Crimerate vs Residual plot")
abline(a=0, b=0, col='red')
```

plot for residual and distance variable
```{r}
plot(train1.data$Distance,rstandard(model2),main="Distance vs Residual plot")
abline(a=0, b=0, col='red')
```

plot for residual and floors variable
```{r}
plot(train1.data$floors,rstandard(model2),main="Floor vs Residual plot")
abline(a=0, b=0, col='red')
```

##Influential Points##

```{r}
cutoff <- 4/((nrow(train1.data)-length(model3$coefficients)-2)) 
plot(model3, which=4, cook.levels=cutoff)
```

Removing influential points
```{r}
fulldata2 <- fulldata[- 19983,]
fulldata2 <- fulldata[- 12699,]
fulldata2 <- fulldata[- 8845,]
data2<-fulldata2[c(2,3,13,14,15,18,19,20,21,22,31,33,34,35)]
```

Building the model again using model3 after removing influentail points
```{r}
selectdata2 = sample(1:nrow(data2),0.80*nrow(data2));
train2.data = data2[selectdata2,]; 
test2.data = data2[-selectdata2,];
```
```{r}
model5 <- lm(formula=log(price)~bathrooms+grade+floors+
           newsqft_living+crimerate+Distance,data=train2.data)
summary(model5)
```

predicting the price for test data using model5
```{r}
y5<-exp(predict.glm(model5,test2.data))
```
Calculating rmse for predicted and observed values using model5
```{r}
y=test2.data[,3]
rmse5= sqrt((y-y5)%*%(y-y5))/nrow(test2.data)
rmse5
```

Residual  Plots
model5
Residual plot normality

```{r}
qqnorm(rstandard(model5))
qqline(rstandard(model5,col="red"))
```

plot between residual and predicted
```{r}
plot(fitted(model5),rstandard(model5),main="predicted vs residual plot")
abline(a=0, b=0, col='red')
```


##########################Hypothesis Testing the effect of distance on price#####
##Two Sample Z test for Distance##

Calculating price persqft 
```{r}
fulldata$pricepersqft <- fulldata$price/fulldata$newsqft_living
names(fulldata)
```
Analysing distance
```{r}
summary(fulldata$Distance)
boxplot(fulldata$Distance)
names(fulldata)
```

Storing distance and price in a new dataset
```{r}
hypdistance<- fulldata[c(2,36)]
```

First sample with zipcodes less then median distance from Amazon head quartes
```{r}
x<- subset(hypdistance,Distance <= 15.85)
```

Second sample with zipcodes gretera then median distance from Amazon head quartes
```{r}
y<- subset(hypdistance,Distance > 15.85)
```
Assiging pricepersqft of two samples to x and y varaibles
```{r}
x <- x$pricepersqft
y<- y$pricepersqft
```
Hypothesis for distance
Null Hypothesis -No Average differnece in house price based on Distance
Alternate Hypothesis -There is Average difference in house pricesummary

Standard deviation of X
```{r}
sd(x)
```

Standard deviation of Y
```{r}
sd(y)
```

Two sample Ztest
```{r}
t.test(x,sigma.x=121.73 ,y,sigma.y=76.98 ,alternative="two.sided" ,conf.level=0.95)
```


##Hypothesis Testing the effect of crime on price##
Two Sample Z test for crime
Calculating price persqft 
```{r}
fulldata$pricepersqft <- fulldata$price/fulldata$newsqft_living
names(fulldata)
```

Analysing crimerate
```{r}
summary(fulldata$crimerate)
boxplot(fulldata$crimerate)
names(fulldata)
```
Storing price per square feet and crime in a data frame
```{r}
hypdistance<- fulldata[c(35,36)]
```
First sample with zipcodes having  crime rates lesser than median crime rate
```{r}
x<- subset(hypdistance,crimerate <= 0.004)
```
Second sample with zipcodes having  crime rates lesser than median crime rate
```{r}
y<- subset(hypdistance,crimerate > 0.004)
```

Assigning the price per squaree feet of two samples to x and y
```{r}
x<- x$pricepersqft
y<- y$pricepersqft
```

Hypothesis for crime
Null Hypothesis -Crime has no effect on average price per square feet
Alternate Hypothesis- Crime is contributing on average price per square feet

Standard deviation of X
```{r}
sd(x)
```

Standard deviation of Y
```{r}
sd(y)
```

Two sample Ztest
```{r}
t.test(x,sigma.x=127.7317 ,y,sigma.y=78.9124 ,conf.level=0.95)
```







